---
title: 'Assignment 2: Data Ingest'
author: "Monique Schafer"
date: "5/9/2017"
output:
  pdf_document: default
  html_document: default
---

###Part 1: Importing and Tidying Data
#####Loading neccessary packages and importing data
######Note: Columns were assigned appropriate type in this step
```{r gazateer, include=TRUE, echo=TRUE}
library(tidyverse)
gaz_raw <- read_delim("CA_Features_20170401.txt", delim = "|", col_types = cols(
  FEATURE_ID = col_character(),
  DATE_CREATED = col_date(format = "%m/%d/%Y"),
  DATE_EDITED = col_date(format = "%m/%d/%Y")
))
```

#####Selecting columns
```{r gaz, include=TRUE, echo=TRUE}
gaz <- select(gaz_raw, FEATURE_ID,FEATURE_NAME, FEATURE_CLASS, STATE_ALPHA, COUNTY_NAME, PRIM_LAT_DEC, PRIM_LONG_DEC,SOURCE_LAT_DEC,SOURCE_LONG_DEC,ELEV_IN_M,MAP_NAME, DATE_CREATED, DATE_EDITED)

gaz[, 6:7][gaz[, 6:7] == 0] <- NA

####Removing Unknown Rows and Selecting For California Features Only
gaz <- gaz%>% filter(STATE_ALPHA == "CA") %>% 
  filter(!is.na(PRIM_LAT_DEC)) %>%
  filter(!is.na(PRIM_LONG_DEC))

write_csv(gaz,"gaz.csv")
```

###Part 2: Analyzing the Gazateer Data
##### What is the most-frequently-occurring feature name? What is the least-frequently-occuring feature class?
```{r gazfeature, include=TRUE, echo=TRUE}
feature_class <- count(gaz, vars = FEATURE_CLASS)
print(feature_class)
write.csv(feature_class, "features.csv")
```
######The most-frequently-occurring feature class is "locale", the least-frequently-occurring feature class is "isthmus" and "sea".

#####What is the approximate center point of each county?
``` {r gazcenterpoint, include=TRUE, echo=TRUE}
gaz_countygeo <- gaz %>%
  group_by(COUNTY_NAME) %>%
  summarise(county_minlat = min(PRIM_LAT_DEC, na.rm = TRUE), 
  county_maxlat = max(PRIM_LAT_DEC, na.rm = TRUE), county_minlong = max(PRIM_LONG_DEC, na.rm = TRUE), county_maxlong = min(PRIM_LONG_DEC, na.rm = TRUE)
  )

gaz_countymid <- transmute(gaz_countygeo,
                        COUNTY_NAME = COUNTY_NAME,
                        MID_LAT = (county_maxlat + county_minlat)/2,
                        MID_LONG = (county_maxlong + county_minlong)/2
)

print(gaz_countymid[,1:3], caption = "Latitude and Longitude Midpoints") 
```
######The midpoints for each county are shown in the table above.

#####What are the fractions of the total number of features in each county that are natural? Man-made?
```{r gazmanmade, include=TRUE, echo=TRUE}
features_natman <- read_csv("features_natman.csv")
all_features_natman <- left_join(features_natman, gaz, by = "FEATURE_CLASS")

count_natural <- count(all_features_natman, vars = Nat_Man, by = COUNTY_NAME)
count_natural_tidy <- spread(count_natural, key = vars, value = n)
prop_table_county <- mutate(count_natural_tidy, 
                            prop_natural = natural / (manmade+natural),
                            prop_manmade = manmade / (natural+manmade)
)
print(prop_table_county)

```